Data preparation report

### 2023-05-26 00_dataPrepReport.Rmd
Added code to remove in-cage dates

The resulting data has dropped from **`r nrow(dat2)`** to
**`r nrow(datAnnot)`** rows, a
**`r round((nrow(dat2)-nrow(datAnnot))*100/nrow(dat2), 3)`% decrease**.
It has `r ncol(datAnnot)` columns and
**`r length(unique(datAnnot$Nili_id))`** unique individuals.

@@

datAnnot2 <- datAnnot %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot)
length(unique(datAnnot2$Nili_id)) # 125
```

>>>>

```{r remove-cages}
# Remove cages using Marta's data
# Identify and remove capture dates using Marta's code --------------------
## To identify the capture dates, first we need to classify the roosts (now using the get_roosts() function). Then, if the bird roosted within 500 m of the capture site, it was considered to be captured and that day, and the following day, are excluded from the dataset. This has been validated. Note: this protocol doesn't work for the Carmel because of the position of the roost sites relative to cages. But we're just dealing with southern individuals, so that's okay.
#roosts <- get_roosts_df(df = datAnnot, id = "Nili_id", timestamp = "timestamp", x = "location_long", y = "location_lat", ground_speed = "ground_speed", speed_units = "m/s", quiet = F)
#save(roosts, file = "data/roosts.Rda")
load("data/roosts.Rda")

# Identify the period of time during which the capture sites are open (when we need to do this exclusion)
start.day <- 01
start.month <- 08
end.day <-  30
end.month <- 11
distance <- 500 # distance, in meters, to calculate from the cage

# Load the information about the capture sites
sites <- read.csv("data/capture_sites.csv")

# Subset the roost dataset with the start and end dates for the capture period
sub.roosts <- roosts %>%
  mutate(start_date = as.Date(paste(start.day, start.month, lubridate::year(date), sep="-"),
                              format="%d-%m-%Y"),
         end_date = as.Date(paste(end.day, end.month, lubridate::year(date), sep="-"),
                            format="%d-%m-%Y")) %>%
  filter(date >= start_date & date <= end_date)

unique(lubridate::month(sub.roosts$date)) # now only includes the fall months, which is the capture season.

# then we need to calculate the roost distance to each of the capture cages. if it is less than 500m, keep that line
crds <- matrix(c(sub.roosts$location_long, sub.roosts$location_lat), 
               nrow = nrow(sub.roosts), ncol = 2) # get roost locations as simple lat/long coordinates

DistanceMat <- matrix(ncol = nrow(sites), nrow = nrow(crds)) 
colnames(DistanceMat) <- unique(sites$name)

for(i in 1:nrow(crds)){ # for each roost point...
  DistanceMat[i,] <- round(geosphere::distm(crds[i,], sites[,c(3,2)]), 2) # calculate distance using geosphere::distm
  print(i)
}

ClosestCaptureSite <- colnames(DistanceMat)[apply(DistanceMat,1,which.min)] # ID of closest capture site
ClosestCaptureDist <- apply(DistanceMat,1,min) # Distance from closest capture site

sub.roosts <- cbind(sub.roosts, ClosestCaptureSite, ClosestCaptureDist) 
sub.roosts$Captured <- ifelse(sub.roosts$ClosestCaptureDist <= distance, "Yes", "No")

sub.captured <- subset(sub.roosts, Captured == "Yes") # get list of dates which the bird was inside the cage

sub.captured.dates <- subset(sub.captured, 
                             select = c("Nili_id",
                                        "date", 
                                        "ClosestCaptureSite",
                                        "ClosestCaptureDist", 
                                        "Captured"))

## This, however, does not work for the Carmel, because the roosts are within 500m of the cage and very often the birds roost on top of the cage without actually being inside it. So for the Carmel captures, we will use another protocol: if the birds were sleeping within 50m of the cage and it was a capture day (or 3 days before the release day), we consider the birds were captured and we remove the data

sub.captured.no.carmel <- subset(sub.captured.dates, ClosestCaptureSite != "Carmel")
sub.captured.carmel <- subset(sub.captured.dates, ClosestCaptureSite == "Carmel")

AllCarmelDates <- read.csv("data/all_captures_carmel_2010-2021.csv")
AllCarmelDates$Date <- as.Date(AllCarmelDates$Date, format = "%d/%m/%Y")

AllCarmelDates.1 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-1)))
AllCarmelDates.2 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-2)))
AllCarmelDates.3 <- data.frame(Date = as.Date(paste(AllCarmelDates$Date-3)))

AllCarmelDates.all <- rbind(AllCarmelDates, AllCarmelDates.1, AllCarmelDates.2, AllCarmelDates.3)

sub.captured.carmel <- sub.captured.carmel %>%
  mutate(known_capture = ifelse(date %in% AllCarmelDates.all$Date, 1, 0),
         Captured = ifelse(known_capture == 1 & ClosestCaptureDist <= 50, "yes", "no")) %>%
  filter(Captured == "yes") %>%
  dplyr::select(-c(known_capture))

names(sub.captured.no.carmel)
names(sub.captured.carmel)

sub.captured.dates <- rbind(sub.captured.no.carmel, sub.captured.carmel)

# We also need to exclude the day after the bird was captured
sub.captured.dates.1 <- sub.captured.dates
sub.captured.dates.1$date <- sub.captured.dates.1$date+1

sub.captured.dates <- rbind(sub.captured.dates, sub.captured.dates.1)
sub.captured.dates <- sub.captured.dates %>% 
  dplyr::distinct(Nili_id, date, .keep_all = T)

# It all looks ok, so we can subset the dataset to exclude the capture dates
datAnnot_noCaptures <- datAnnot %>%
  left_join(sub.captured.dates, by = c("Nili_id", "dateOnly" = "date"))
nrow(datAnnot) == nrow(datAnnot_noCaptures) # should be TRUE. NOW we can filter.
datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::filter(Captured != "Yes"|is.na(Captured)) # remove the individual*days when they were captured

datAnnot_noCaptures <- datAnnot_noCaptures %>%
  dplyr::select(-c("ClosestCaptureSite", "ClosestCaptureDist", "Captured"))

# How many rows and individuals did we remove?
before <- nrow(datAnnot)
after <- nrow(datAnnot_noCaptures)
(propChange <- (before-after)/before) # approx. 1% of data removed.
length(unique(datAnnot$Nili_id))
length(unique(datAnnot_noCaptures$Nili_id)) # no change in number of individuals! As expected.
```

The resulting data has dropped from **`r nrow(datAnnot_noCaptures)`** to
**`r nrow(datAnnot)`** rows, a
**`r round((nrow(datAnnot_noCaptures)-nrow(datAnnot))*100/nrow(datAnnot_noCaptures), 3)`% decrease**.
It has `r ncol(datAnnot)` columns and
**`r length(unique(datAnnot$Nili_id))`** unique individuals.

@@

datAnnot2 <- datAnnot_noCaptures %>%
  dplyr::select(-c("sex")) %>%
  left_join(as, by = "Nili_id")
nrow(datAnnot2) == nrow(datAnnot)
length(unique(datAnnot2$Nili_id)) # 125
```
